Iterative Reduction: Parsing and Interpreting Code Without Trees

### 2024-12-17 Paper/paper.tex
Cleaned up comments

    Now, consider the expression $1*2+3$. 
    After assigning priorities, we obtain the following expression:
    $\mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty$.
    \yoni{The last sentence is an example of how the assignment
    of priorities feels a bit out of context here. This could improve
    if they become a formal part of the framework, from the very beginning.}
    \ari{Done, thoughts?}
    We now get:
    \begin{multline*}
        \mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt{2,+}_1 \mbt3_\infty \to \mbt{2,+}_1 \mbt3_\infty\\
        \to \mbt{2,+}_1 \mbt3_0 \to \mbt5_0
    \end{multline*}

@@

But this has a major issue: it can't handle things like parentheses.
We could of course handle parentheses by changing how we assign priorities: every time we see an open parentheses we could increase
the priority of each token and then beta-reduce.
But this complicates how we allocate priorities, as we can no longer statically assign priorities using a simple priority function $\pi$ but must instead design a more complicated algorithm in order
to allow for generalization.
\yoni{would u say that currently we can set priorities statically, but with
the idea of increasing priorities we would have to set them dynamically?
If so, I would mention that.}
\ari{Done, thoughts?}
Furthermore the notion introduced in the next section, partial reduction, generalizes well and is used as a basepoint in the
remainder of this paper.

@@

    \yoni{I don't really understand it. For example, what is 
    $(v,f\mapsto(v,f))$? Perhaps introducing some more definitions
    and details can help clarify?
    }
    \ari{Explained, is this good?}
    \yoni{What happened to the squares around the tokens in this example?}

@@

    \yoni{Here we are missing a formal assignment of priorities.}
    \ari{I added a few sentences explaining this.}

@@

\yoni{I don't think this is very standard. Perhaps you can define
 it inductively: Every element of $\Sigma^0$ is in $T_{\Sigma}$,
  and for eac $n$, if $t_i\in T_{\Sigma}$ for each $1\leq i \leq n$
  then $f t_1 \ldots t_n \in T_{\Sigma}$ for each $f\in\Sigma^{n}$,
  and then for each $g\in\Sigma^{\infty}$ and any $n$, similarly...
Do we actually need $\Sigma^{\infty}$ BTW? }
\ari{Done.
And yeah, $\Sigma^\infty$ is important since products have a variable number of parameters.}

@@

For readability we can put and omit parentheses and commas where convenient, but importantly these sets have the \textit{unique reconstruction property}:
it is possible to compute their construction from the string itself (i.e. a string cannot be constructed from two different patterns).
This property is not hard to prove.
\yoni{If I understand correctly you are assuming this property, right? If so, please say so.}
\ari{No, by the definition of these sets, they have this property.}

@@

    If $\Sigma^{0}=\{\NUM\}$, $\Sigma^{1}=\{\LIST\}$, and $\Sigma^{\infty}=\{\PRODUCT\}$, and $\Sigma=\Sigma^0\cup\Sigma^1\cup\Sigma^\infty$, and $\c A=\mathcal A^1=\set{\OP}$, 
    \yoni{plz specify the "arity" of $\OP$.}
    \ari{Done}
    then
    \[ \NUM,\LIST\ \NUM,\PRODUCT(\LIST\ \NUM\ \NUM)\in\terms,\qquad \OP\ \LIST\ \NUM\in\terma \]
    \yoni{I think it is better to put a space between these, especially
    in $\LIST\NUM\NUM$.}
    \ari{Done}

@@

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terma$, 
        \yoni{Should the last $\terms$ be $\terma$?}
        \ari{Yeah}
    the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\expi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before: $\sigma$ is the term (printable or internal), $n$ the priority, and $v$ the value.
In the case that $\sigma$ is a printable term, we still use this notation even though the value does not exist: so we just take it to be empty.
\yoni{But elements of $\oe$ seems like pairs to me, not triplets.}
\ari{There was a slight mistake in the definition of $\exp$.}
% for printable terms we act as if the value is empty.
\yoni{Please clarify what does acting as if... means.}
\ari{Rewritten}

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$ as before: $\pi(\sigma^1\cdots\sigma^n)=\sigma^1_{\rho_1}\cdots\sigma^n_{\rho_n}$ where $\rho_i=\pi(\sigma^i)$.
\yoni{please show how.}
\ari{Done}

>>>>

    Now, consider the expression $1*2+3$. 
    After assigning priorities, we obtain the following expression:
    $\mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty$.
    We now get:
    \begin{multline*}
        \mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt{2,+}_1 \mbt3_\infty \to \mbt{2,+}_1 \mbt3_\infty\\
        \to \mbt{2,+}_1 \mbt3_0 \to \mbt5_0
    \end{multline*}

@@

But this has a major issue: it can't handle things like parentheses.
We could of course handle parentheses by changing how we assign priorities: every time we see an open parentheses we could increase
the priority of each token and then beta-reduce.
But this complicates how we allocate priorities, as we can no longer statically assign priorities using a simple priority function $\pi$ but must instead design a more complicated algorithm in order
to allow for generalization.
Furthermore the notion introduced in the next section, partial reduction, generalizes well and is used as a basepoint in the
remainder of this paper.

@@

@@

@@

@@

For readability we can put and omit parentheses and commas where convenient, but importantly these sets have the \textit{unique reconstruction property}:
it is possible to compute their construction from the string itself (i.e. a string cannot be constructed from two different patterns).
This property is not hard to prove.

@@

    If $\Sigma^{0}=\{\NUM\}$, $\Sigma^{1}=\{\LIST\}$, and $\Sigma^{\infty}=\{\PRODUCT\}$, and $\Sigma=\Sigma^0\cup\Sigma^1\cup\Sigma^\infty$, and $\c A=\mathcal A^1=\set{\OP}$, 
    then
    \[ \NUM,\LIST\ \NUM,\PRODUCT(\LIST\ \NUM\ \NUM)\in\terms,\qquad \OP\ \LIST\ \NUM\in\terma \]

@@

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terma$, the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\expi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before: $\sigma$ is the term (printable or internal), $n$ the priority, and $v$ the value.
In the case that $\sigma$ is a printable term, we still use this notation even though the value does not exist: so we just take it to be empty.

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$ as before: $\pi(\sigma^1\cdots\sigma^n)=\sigma^1_{\rho_1}\cdots\sigma^n_{\rho_n}$ where $\rho_i=\pi(\sigma^i)$.
